Pillar 1 - Sequential Pattern Mining

### 2019-01-30 Pillar 1 - Sequential Pattern Mining.Rmd
finish cleaning ready for KNIT

```{r library, message = F}
library(arulesSequences)

@@

### Adding not taken courses
```{r}
d_transactions_not_taken <- expand.grid(

@@

```
###Expanded Transcripts
```{r}

d_transcript_expanded <- d_transcript %>%
  filter(!is.na(Grade), Grade>=0) %>%
  mutate(Round_grade = paste("g", round(Grade,0),sep=""),
         Values = T) %>%
  spread(key = "Round_grade", value = Values) %>%
  mutate(g1  = g0|g1,
         g2  = g1|g2,
         g3  = g2|g3,
         g4  = g3|g4,
         g5  = g4|g5,
         g6  = g5|g6,
         g7  = g6|g7,
         g8  = g7|g8,
         g9  = g8|g9,
         g10 = g9|g10
         ) %>%
  gather(key="Gr_string", value="Gr_bool", g0, g1, g2, g3, g4,g5,g6,g7,g8,g9,g10, na.rm=T) %>%
  select(-Gr_bool )%>%
  separate(Gr_string, into=c("Trash", "Grade_new"), sep="g")%>%
  select(-Trash) %>%
  mutate(Grade= as.numeric(Grade_new))%>%
  select(-Grade_new)

#changing to transactions
d_transactions_expanded <- d_course %>%

  # Join with transcripts.
  select(- Period) %>%
  inner_join(d_transcript_expanded, by = "Course ID") %>%

    item_PF = paste(`Course ID`, PF, sep = "_"),
    item_HL = paste(`Course ID`, HL, sep = "_")
    ) %>%
  rename(
    item = `Course ID`
  )

rm(d_transcript_expanded)
```
### Transactions for Apriori
```{r making transactions}

@@

@@

### Sequences for CSPADE
```{r}

@@

@@

## Mining Rules
### Apriori Algorithm
```{r}

@@

#
# Save association rules
save(AR_taken, AR_PF, AR_HL, AR_not_taken,
     file = "App/AR.RDATA")

rm(clean_AR, my_apriori,
   transactions, transactions_PF, transactions_HL,
   course_id_fail, course_id_low,
   AR_taken, AR_PF, AR_HL, AR_not_taken)

@@

### CSPADE Algorithm
```{r}
# Transform rules from ruleInduction into a readable data frame
n_students <- length(unique(d_transactions$sequenceID))

@@

```
We apply cspade followed by rule induction
```{r}
# Run the cspade algorithm with desired parameters. Then run the ruleInduction algorithm with desired parameters.
my_SR<- function(data){
 data %>%
  cspade(
    parameter = list(
      support = 0, # min support of a sequence
      maxsize = 1, # max number of items of an element of a sequence
      maxlen  = 2, # max number of element of a sequence
      mingap  = 1, # min time difference between consecutive element of a sequence
      maxgap  = 1e4 # max time difference between consecutive element of a sequence
      #maxwin  = 1e4  # max time difference between any two elements of a sequence # WARNING: 'maxwin' disabled
      ),
    control = list(
      verbose = FALSE
      )
    ) %>%
    ruleInduction(
      confidence = 0, #we need to play with the parameters.
      control    = list(verbose = FALSE)
      ) %>%
    clean_SR
}

@@

#Saving sequence rules
SR <- list(
    SR = SR,
    SR_PF = SR_PF,
    SR_HL = SR_HL
    )

@@

#clean unnecessary objects
rm(clean_SR, my_SR, 
   sequences, sequences_PF, sequences_HL, 
   n_students,
   SR, SR_PF, SR_HL)

>>>>

```{r library, message = FALSE, warning = FALSE}
library(arulesSequences)

@@

### Adding not taken courses
```{r transactions not taken}
d_transactions_not_taken <- expand.grid(

@@

```
###Expanded Transcripts
```{r transactions expanded}

d_transactions_expanded <- d_transactions %>%
  mutate(Round_grade = round(Grade, 0),
         Values = T) %>%
  spread(key = "Round_grade", value = Values) %>%
  mutate(`1`  = `0`|`1`,
         `2`  = `1`|`2`,
         `3`  = `2`|`3`,
         `4`  = `3`|`4`,
         `5`  = `4`|`5`,
         `6`  = `5`|`6`,
         `7`  = `6`|`7`,
         `8`  = `7`|`8`,
         `9`  = `8`|`9`,
         `10`  = `9`|`10`
         ) %>%
  gather(key = "Grade_round", value = "Gr_bool", `0`, `1`, `2`, `3`, `4`, `5`, `6`, `7`, `8`, `9`, `10`, na.rm = TRUE, convert = TRUE) %>%
  select(-Gr_bool) %>%
  arrange(
    sequenceID,
    item
  ) %>%

  # Identifying item
  mutate(
    PF = case_when( Grade_round <  pass_grade ~ "fail",
                    Grade_round >= pass_grade ~ "pass"),
    HL = case_when( Grade_round <  high_grade ~ "low",
                    Grade_round >= high_grade ~ "high"),

    item_PF = paste(item, PF, sep = "_"),
    item_HL = paste(item, HL, sep = "_")
    )
```
### Transactions for Apriori
```{r making transactions}

@@

transactions_expanded         <- make_transaction(data = d_transactions_expanded, item = item)
transactions_expanded_PF      <- make_transaction(data = d_transactions_expanded, item = item_PF)
transactions_expanded_HL      <- make_transaction(data = d_transactions_expanded, item = item_HL)

@@

### Sequences for CSPADE
```{r making sequences}

@@

sequences_expanded     <- make_sequence(data = d_transactions_expanded, item = item)
sequences_expanded_PF  <- make_sequence(data = d_transactions_expanded, item = item_PF)
sequences_expanded_HL  <- make_sequence(data = d_transactions_expanded, item = item_HL)

@@

## Mining Rules
### Apriori Algorithm
```{r function clean AR}

@@

AR_expanded <- my_apriori(transactions_expanded)
# creating vector of fail course
course_id_fail <- d_transactions_expanded %>%
  filter(PF == "fail") %>%
  distinct(`item_PF`)

AR_expanded_PF <- my_apriori(transactions_expanded_PF,
                    appearance = list(both = course_id_fail$`item_PF`))

# creating vector of low score course
course_id_low <- d_transactions_expanded %>%
  filter(HL == "low") %>%
  distinct(`item_HL`)

AR_expanded_HL <- my_apriori(transactions_expanded_HL,
                    appearance = list(both = course_id_low$`item_HL`))

#
# Save association rules
save(AR_taken, AR_PF, AR_HL, AR_not_taken, 
     AR_expanded, AR_expanded_PF, AR_expanded_HL,
     file = "App/AR.RDATA")

rm(clean_AR, my_apriori,
   transactions, transactions_PF, transactions_HL,
   course_id_fail, course_id_low,
   AR_taken, AR_PF, AR_HL, AR_not_taken, AR_expanded, AR_expanded_PF, AR_expanded_HL)

@@

### CSPADE Algorithm
```{r function clean SR}
# Transform rules from ruleInduction into a readable data frame
n_students <- length(unique(d_transactions$sequenceID))

@@

```
We apply cspade followed by rule induction
```{r cspade and rule induction, cache = TRUE}
# Run the cspade algorithm with desired parameters. Then run the ruleInduction algorithm with desired parameters.
my_SR<- function(data){
 data %>%
  cspade(
    parameter = list(
      support = 0, # min support of a sequence
      maxsize = 1, # max number of items of an element of a sequence
      maxlen  = 2, # max number of element of a sequence
      mingap  = 1, # min time difference between consecutive element of a sequence
      maxgap  = 1e4 # max time difference between consecutive element of a sequence
      #maxwin  = 1e4  # max time difference between any two elements of a sequence # WARNING: 'maxwin' disabled
      ),
    control = list(
      verbose = FALSE
      )
    ) %>%
    ruleInduction(
      confidence = 0, #we need to play with the parameters.
      control    = list(verbose = FALSE)
      ) %>%
    clean_SR
}

@@

SR_expanded   <- my_SR(data = sequences_expanded   )
SR_expanded_PF <- my_SR(data = sequences_expanded_PF)
SR_expanded_HL <- my_SR(data = sequences_expanded_HL)
#Saving sequence rules
SR <- list(
    SR = SR,
    SR_PF = SR_PF,
    SR_HL = SR_HL,

    SR_expanded    = SR_expanded,
    SR_expanded_PF = SR_expanded_PF,
    SR_expanded_HL = SR_expanded_HL
    )

@@

#clean unnecessary objects
rm(clean_SR, my_SR, 
   sequences, sequences_PF, sequences_HL,
   sequences_expanded, sequences_expanded_PF, sequences_expanded_HL,
   n_students,
   SR, SR_PF, SR_HL,
   SR_expanded, SR_expanded_PF, SR_expanded_HL)
